 

### 2020-12-25 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

While forest microclimates are insulated from the extremes of macroclimate warming, they are still experiencing warming. Temperature is directly linked with plant metabolic processes and growth, thus even around 1 degree of warming can lead to changes in plant species composition and growth [Bertrand et al.2011] This is leading to thermophilization in plant communities and reshuffling of species acclimated to cooler [mean annual] temperatures[@zellweger_forest_2020; @zellweger_seasonal_2019; @defrenneMicroclimateModeratesPlant2013; @suggittExtinctionRiskClimate2018; @scheffersMicrohabitatsReduceAnimal2013]. 
Microclimate warming has implications for forest regeneration[@vonarxSpatiotemporalEffectsForest2012a] as temperature driven photosynthesis limitation is leading to reduced tree growth rates, driving a global trend towards disequillibrium in forests with younger trees [@mcdowellMechanismsPlantSurvival2008].
Reductions in canopy cover can lead to non-linear threshold responses, causing dramatic shifts from one forest state to another (e.g., transition from forest to savanna-like vegetation in tropical forest regions through "savannization"), with energy balance impacts and implications for forest-atmosphere interactions (Stark et al. 2020). Each of these changes have potential feedbacks to climate change.

@@

>>>>

While forest microclimates are insulated from the extremes of macroclimate warming, they are still experiencing warming. Temperature is directly linked with plant metabolic processes and growth, thus even around 1 degree of warming can lead to changes in plant species composition and growth [Bertrand et al.2011]. This is leading to thermophilization in plant communities and reshuffling of species acclimated to cooler [mean annual] temperatures[@zellweger_forest_2020; @zellweger_seasonal_2019; @defrenneMicroclimateModeratesPlant2013; @suggittExtinctionRiskClimate2018; @scheffersMicrohabitatsReduceAnimal2013]. Considering frequent occurences of drought, extensive analysis also shows negetive correlation between species shade tolerance and drought tolerance globally [Niinemetes and Valladares] which may have further impacts on understory species ontogeny with rising temperatures. 
Microclimate warming has implications for forest regeneration[@vonarxSpatiotemporalEffectsForest2012a] as temperature driven photosynthesis limitation is leading to reduced tree growth rates, driving a global trend towards disequillibrium in forests with younger trees [@mcdowellMechanismsPlantSurvival2008].
Reductions in canopy cover can lead to non-linear threshold responses, causing dramatic shifts from one forest state to another (e.g., transition from forest to savanna-like vegetation in tropical forest regions through "savannization"), with energy balance impacts and implications for forest-atmosphere interactions (Stark et al. 2020). Each of these changes have potential feedbacks to climate change.

@@

Cateogorization of sun and shade tolerance from Niinemetes and Valladares, 2006: "The five-level scale used for shade tolerance (1,
very intolerant; 2, intolerant; 3, moderately tolerant; 4,tolerant; 5, very tolerant) corresponds approximately to the following light availabilities expressed as percentage of full sunlight: 1,>50%; 2, 25-50%; 3, 10-25%; 4, 5-10%; 5, 2-5%"
